Clause size reduction with all-UIP Learning

### 2020-02-24 paper.tex
fix typo, update proof reduction results with 30000 proof checking time

The LBD of the learnt clause $C_L$ is the number of different decision
levels in it:
$\LBD(C_L)=\left|\big\{\dlevel{l}\,|\,l \in C_L\big\}\right|$
\cite{DBLP:conf/ijcai/AudemardS09}. Empirically LBD is a successful
predictor of clause usefulness: clauses with lower LBD tend to be more
useful. As noted in \cite{DBLP:conf/ijcai/AudemardS09}, from the
initial falsified clause $C_I$ the 1-UIP scheme will produce a clause
$C_L$ whose LBD is minimum among all asserting clauses that can be
learnt from $C_I$. If $C'$ is a trail resolvant of $C$ and a reason
clause $\reason{l}$, then $\LBD(C') \geq \LBD(C)$ since $\reason{l}$
must contain at least one other literal with the same decision level
as $l$ and might contain literals with decision levels not in
$C$. That is, each trail resolution step can only increase the \LBD of
the learnt clause. Hence, the 1-UIP scheme yields an asserting clause
with minimum LBD as it performs the minimum number of trail resolutions
required to generate an asserting clause.

@@

\subsubsection{$\allUipAct$: Clauses with Active
  Variables.} \label{sec:active} $\allUip$ learning might introduce
literals with low variable activity into $C_i$. Low activity variables
are variables that have had low recent participation in clause
learning. Hence, clauses with variables of low activity might not be
as currently useful to the solver. Our variant $\allUipAct$, shown as
the optional line~\ref{ln:activity} in Algorithm~\ref{alg:i-uip},
computes the average variable activity of the newly produced all-UIP
clause $C_i$ and the original 1-UIP clause $C_1$. The new clause $C_i$
will be returned only if it is both smaller and has higher average
variable activity than the original 1-UIP clause. There are, of
course, generalizations of this approach where one has a weighted
tradeoff between these factors that allows preferring the new clause
when it has large gains in one metric even though it has small losses
in the other. We did not, however, experiment with such
generalizations.

@@

\subsubsection{Reduced Proof sizes with $\allUip$.}
A learning scheme that yields smaller clauses (lemmas) might also
construct smaller causal proofs. For UNSAT instances, we additionally
compare the size of the optimized DRAT proof from $\allUipPure$,
$\allUipMin$ and 1-UIP schemes. We used the DART-trim tool
\cite{DBLP:conf/sat/WetzlerHH14} with a 10000 second timeout to check and
optimize every DRAT proof once \footnote{Applying DART-trim multiple
  times can further reduce the proof size until a fix-point. However,
  the full optimization is time consuming and was impractical for our
  experiments.}.

The average optimized DRAT proof from $\allUipMin$ and $\allUipPure$
are 413.2MB and 487.2MB, respectively. Both sizes are significantly
smaller than the average optimized proof size from 1-UIP, 613.9MB.
The average proof size reduction per instance for $\allUipMin$ and
$\allUipPure$ was 17.18\% and 6.9\% compared with 1-UIP, which roughly
correlate with our observed clause size observation in
Fig~\ref{fig:len_pdf}.

>>>>

The LBD of the learnt clause $C_L$ is the number of different decision
levels in it:
$\LBD(C_L)=\left|\big\{\dlevel{l}\,|\,l \in C_L\big\}\right|$
\cite{DBLP:conf/ijcai/AudemardS09}. Empirically LBD is a successful
predictor of clause usefulness: clauses with lower LBD tend to be more
useful. As noted in \cite{DBLP:conf/ijcai/AudemardS09}, from the
initial falsified clause $C_I$ the 1-UIP scheme will produce a clause
$C_L$ whose LBD is minimum among all asserting clauses that can be
learnt from $C_I$. If $C'$ is a trail resolvent of $C$ and a reason
clause $\reason{l}$, then $\LBD(C') \geq \LBD(C)$ since $\reason{l}$
must contain at least one other literal with the same decision level
as $l$ and might contain literals with decision levels not in
$C$. That is, each trail resolution step can only increase the \LBD of
the learnt clause. Hence, the 1-UIP scheme yields an asserting clause
with minimum LBD as it performs the minimum number of trail resolutions
required to generate an asserting clause.

@@

\subsubsection{$\allUipAct$: Clauses with Active
  Variables.} \label{sec:active} $\allUip$ learning might introduce
literals with low variable activity into $C_i$. Low activity variables
are variables that have had low recent participation in clause
learning. Hence, clauses with variables of low activity might not be
as currently useful to the solver. Our variant $\allUipAct$, shown as
the optional line~\ref{ln:activity} in Algorithm~\ref{alg:i-uip},
computes the average variable activity of the newly produced all-UIP
clause $C_i$ and the original 1-UIP clause $C_1$. The new clause $C_i$
will be returned only if it is both smaller and has higher average
variable activity than the original 1-UIP clause. There are, of
course, generalizations of this approach where one has a weighted
trade-off between these factors that allows preferring the new clause
when it has large gains in one metric even though it has small losses
in the other. We did not, however, experiment with such
generalizations.

@@

\subsubsection{Reduced Proof sizes with $\allUip$.}
A learning scheme that yields smaller clauses (lemmas) might also
construct smaller causal proofs. For 88 UNSAT instances solved mutually
by $\allUipPure$, $\allUipMin$ and 1-UIP schemes, we additionally
compare the size of the optimized DRAT proof from the three learning schemes. 
We used the DART-trim tool
\cite{DBLP:conf/sat/WetzlerHH14} with a 30000 second timeout to check and
optimize every DRAT proof once \footnote{Applying DART-trim multiple
  times can further reduce the proof size until a fix-point. However,
  the full optimization is time consuming and was impractical for our
  experiments.}.

The average optimized DRAT proof from $\allUipMin$ and $\allUipPure$
are 556.6MB and 654.4MB, respectively. Both sizes are significantly
smaller than the average optimized proof size from 1-UIP, 824.9MB.
The average proof size reduction per instance for $\allUipMin$ and
$\allUipPure$ was 16.5\% and 4.6\% compared with 1-UIP, which roughly
correlate with our observed clause size observation in
Fig~\ref{fig:len_pdf}.
